\center Towards Globally Color-Dual All-Loop Integrands

### 2023-08-17 notes/2loopNLSM.tex

\tikzset{graviton/.style={decorate, decoration={snake, amplitude=.4mm, segment length=1.5mm, pre length=.5mm, post length=.5mm}, double}}
\tikzfeynmanset{compat=1.1.0}
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\usetikzlibrary{external}  
%          %% Load the `external` library
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\tikzexternalize[                     %% Activate externalization
%  prefix=images/,                    %% Avoid cluttering the directory
% system call={                       %% Use lualatex in system call
%    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
%  },
%]
\tikzfeynmanset{
  fermion2/.style={
    /tikz/postaction={
      /tikz/decoration={
        markings,
        mark=at position 0.7 with {
          \arrow{>[length=6pt, width=7pt]};
        },
      },
      /tikz/decorate=true,
    },
  },
}

@@

\newcommand{\simplebox}{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-1,1){2};
\vertex (a2) at (1,-1){4};
\vertex (a3) at (1,1){3};
\vertex (a4) at (-1,-1){1};
\vertex (mid3) at (.35,.35);
\vertex (mid4) at (.35,-.35);
\vertex (mid5) at (-.35,.35);
\vertex (mid6) at (-.35,-.35);
\diagram{(mid4) --[ultra thick,nhpRed,quarter right](mid3),
(mid5) --[ultra thick,nhpRed,quarter right](mid6),
(mid3) --[ultra thick,nhpRed,quarter right](mid5),
(mid6) --[ultra thick,nhpRed,quarter right](mid4),
(mid6) --[ultra thick,fermion2](a4),
(mid3) --[ultra thick,fermion2](a3),
(mid4) --[ultra thick,fermion2](a2),
(mid5) --[ultra thick,fermion2](a1),
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\def\dj#1{{\color{NUpurple}\it \bf JJ: #1}}

@@

\author[1]{Alex Edison,}
\author[1]{James Mangan,}
\author[1]{Nicolas H. Pavao}

\affiliation[1]{Department of Physics and Astronomy, Northwestern
  University, Evanston, Illinois 60208, USA}

\title{Towards Globally Color-Dual All-Loop Integrands}

@@

\langle ab\rangle \equiv p_a^0 p_b^1 - p_a^1 p_b^0 = -\langle ba \rangle 

@@

 From the interaction part of the Lagrangian, we can see that the three-point vertex is (1) functionally symmetric and (2) the kinematics are completely anti-symmetric: 

@@

\cubic{hgrey0}{}{nhpBlue}{}{nhpBlue}{}{nhpBlue} = \langle 12\rangle = \langle 23\rangle = \langle 31\rangle = -\langle 21\rangle

@@

This important property is due to momentum conservation alone and does not need any additional kinematic constraints to satisfy the duality. In addition to anti-symmetry, the cubic graphs will satisfy jacobi off-shell on any internal edge due to the delta function identity of 2D Levi-Civita symbols:

@@

\langle ab \rangle \langle cd\rangle +\langle ac \rangle \langle db\rangle +\langle ad \rangle \langle bc\rangle = \delta^{[a}_c\delta^{b]}_d + \delta^{[a}_d\delta^{c]}_b+ \delta^{[a}_b\delta^{d]}_c=0

@@

Theories with this property are the creme de la creme of color-dual gauge theories as they trivially satisfy color-kinematic algebraic relations off-shell. This allows one to trivially promote the duality from tree-level to all-loop order by dressing each cubic vertex with its own kinematic structure constant.

While the color-dual nature of ZM theory is rather elegant, the loop-level construction introduces complications when applying dimensional regularization. Similar to the difficulties of renormalizing chiral fermions \cite{ 'tHooft}, there is ambiguity in promoting the integrands to formal $D$-dimensional expressions. 
\subsection{YZ-Theory}
\be
\mathcal{L}^{\text{YZ}} =\frac{1}{2} (\partial Y)^2 + (\partial Z)(\partial \bar{Z}) - g f^{abc} \left( \bar{Z}_{\mu\nu}Z^{\mu} Z^\nu + [Y,\partial_\mu Y] Z^\mu \right)
\ee
\be
\Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} = i \varepsilon_3 (p_1-p_2)\qquad \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)
\ee
We begin with a review of the construction of NLSM numerators at tree-level. For this work we extend the tree-level construction of $Y\!Z$ model of Cheung and Hsien \cite{Cheung2016prv}. In this construction, the generators of the kinematic algebra can be expressed as follows
\be\label{eq:FeynmanRuleYYZ}
T^a_{ij}= i \varepsilon_a(p_i-p_j)
\ee
where momentum conservation requires
\be
p_a + p_i + p_j =0
\ee
The kinematic half-ladder diagrams then takes on the following concise definition:
\be
n^{\text{NLSM}}_{(i|a_1a_2...a_n|j)} = (T^{a_1}T^{a_2}\cdots T^{a_n})_{ij}
\ee
Since there are on pole cancelling factors of $s_{ij} = (p_i+p_j)^2$, this definition of the chiral current algebra is manifestly cubic. Thus, the kinematic structure constants defined in terms of these generators are invariant under generalized gauge freedom. They can be defined implicitly below:
\be
[T^a,T^b]_{ij}= F^{a}_{\,b|c}T^c_{ij}
\ee
Given this definition, the Feynman rule associated with kinematic structure constant is simply:
\be\label{eq:FeynmanRuleXZZ}
i F^{a}_{\,b|c} = (\varepsilon_b p_{ab})(\varepsilon_a\bar{\varepsilon}_c) - (\varepsilon_a p_{ab})(\varepsilon_b\bar{\varepsilon}_c) 
\ee
where $p_{ab}=p_a+p_b$ and $\epsilon$ and $\bar{\epsilon}$ are the polarizations of the $Z$-vectors particle and its conjugate field, respectively. The on-shell state sum for the polarization vectors is simply:
\be
\sum_{\text{states}} \varepsilon^{\,\mu}_{(p)}\bar{\varepsilon}^{\,\nu}_{(-p)} = \eta^{\mu\nu}
\ee
Notice that the vector state sum is gauge fixed since the $Y\,Z$ model explicitly chooses Lorenz gauge for the $Z$ particles, $\partial_\mu Z^\mu=0$. To recover NLSM amplitudes from this these kinematic structure constants at tree-level, we simply plug in the following on-shell polarizations for the $Z$ and $\bar{Z}$ particles:
\be\label{eq:onShellZStates}
\varepsilon^\mu_{(p)} = p^\mu \qquad \bar{\varepsilon}^\mu_{(p)} = \frac{q^\mu}{pq}
\ee
where $q^2=0$ is some null reference momentum. With this, we can define tree-level pion scattering in two equivalent ways:
\be
A^{\text{NLSM}} = A(...,Y,...,Y,...) = A(...,\bar{Z},...) 
\ee
where the ellipses denote additional on-shell $Z$-particles. Subject to a particular gauge choice, the kinematic numerators in latter definition for pion scattering is equivalent to those $J$-theory, first written down by Cheung and one of the authors \cite{Cheung:2021zvb}. In this paradigm, the on-shell $\bar{Z}$ state is equivalent to the root leg appearing in the color-dual $J$-theory numerators.

It is instructive to see how both of these constructions produce valid tree-level amplitudes for the pion. First we'll start with two $Y$ particles on legs 1 and 4. Applying the Feynman rules above, and plugging in on-shell states for the $Z$-particles produces the following $s$- and $t$-channel numerators:
\begin{align}
n^{YY}_s &= (T^2T^3)_{14} = s_{12}^2 
\\
 n^{YY}_t &=  F^{3}_{\,2|X}T^X_{14}  = s_{14}(s_{13}-s_{12})
\end{align}
Plugging these numerators into the ordered amplitudes $A(s,t)$ yields the desired result:
\be\label{eq:NLSMYZ4point}
A^{YY}_{(s,t)} = \frac{n^{YY}_s}{s_{12}}+\frac{ n^{YY}_t }{s_{14}} = s_{13}
\ee
Similarly we can do the same for the $Z$ and $\bar{Z}$ configuration. Below we have plugged in the on-shell $Z$-particle states, but have left the $\bar{Z}$ index free. This produces the following numerators:
\begin{align}
n^{\bar{Z}Z}_s &= {}_4(F^{3}F^{2})_{1} =  s_{12}(s_{14}-s_{13})p_2^{\mu_1}+s_{12}^2(p_3-p_4)^{\mu_1}
\\
 n^{\bar{Z}Z}_t &=   {}_2(F^{3}F^{4})_{1}  = s_{14}(s_{12}-s_{13})p_4^{\mu_1}+s_{14}^2(p_3-p_2)^{\mu_1}
\end{align}
where we have defined the short hand notation:
\be
{}_x(F^{a_1}F^{a_2}\cdots F^{a_n})_{y} \equiv F^{a_1}_{\,x|b_2}F^{a_2}_{\,b_2|b_3}\cdots F^{a_n}_{\,b_n|y}
\ee
Similarly, we find the following ordered amplitude when plugging in these numerators:
\be
A^{\bar{Z}Z}_{(s,t)} = \frac{n^{\bar{Z}Z}_s}{s_{12}}+\frac{ n^{\bar{Z}Z}_t }{s_{14}} = -s_{13}(p_2+p_3+p_4)^{\mu_1} = s_{13} \,p_1^{\mu_1}
\ee
Plugging in the on-shell polarizatoin of the conjugate field in \eqn{eq:onShellZStates} produces precisely the desired result of \eqn{eq:NLSMYZ4point}. Indeed, this construction is valid to all orders at tree-level. There are only two possible factorization channels the contribute the each of these amplitudes, the $YY$ cut and the $\bar{Z}Z$ cut :
\begin{align}
A(...,Y,...,Y,...) &\rightarrow A(...,Y,...,Y)A(Y,...,Y,...)
\\
&\rightarrow A(...,Y,...,Y,...,Z)A(\bar{Z},...)
\end{align}
\subsection{J-theory}
\be
\mathcal{L}^{\text{J}} = \frac{1}{2}\partial_\mu J_\nu \bar{J}^{\mu\nu} +g f^{abc} \bar{J}_{\mu\nu}J^\mu J^\nu
\ee
where as above, $\bar{J}_{\mu\nu} = \partial_\mu J_\nu - \partial_\nu J_\mu$. Subject to the Lorenz gauge condition, $\partial_\mu J^\mu =0$, this Lagrangian produces precisely the EOM for J-theory needed for Berends-Giele recursion at tree-level: 
\be
\partial^2 J^\mu - gf^{abc} J^\nu \partial_\nu J^\mu = 0 
\ee
As such, the single vertex in this theory is the same as for YZ theory: 
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)
\ee

\be
\mathcal{L}^{\text{SDYM}} = \frac{1}{2}(\partial \bar{\Phi} )^a(\partial  \Phi)^a +g f^{abc} \bar{\Phi} ^a (\partial_u \Phi)^b (\partial_v \Phi)^c
\ee
\be
(\varepsilon^+_1 k_2) - (\varepsilon^+_2 k_1) = [12]\left(\frac{\langle 2 q\rangle}{\langle 1 q\rangle} + \frac{\langle 1 q\rangle}{\langle 2 q\rangle} \right) \sim [12]
\ee
Now we are prepared to discuss the one-loop case.

\section{One-loop}

\subsection{NLSM -- Nic}

\draftnoteJM{I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.}

\be
\simplebox = (2 f_\pi^{-1})^4 (l_1 k_1)(l_2 k_2)(l_3 k_3)(l_4 k_4)
\ee
where we define sequential loop momenta on the $n$-gon as $l_{i+1}= l_i+k_i$. Given this, the $n$-gon permits a equivalent expression in terms of inverse propagators:
\be
\simplebox =f_\pi^{-4} [12][23][34][41]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_i^L - l_R^2$
\subsection{BELs and integration zeros -- Nic}

At one-loop, weight counting tells us that the $n$-gon master numerator must have $n$ on-shell $Z$-particles. Unitarity requires that there are three distinct contributions from $Y\!Z$ theory -- the first from an off-shell $Y$-loop particle and then two more from different orientations of a $\bar{Z}Z$-loop.  Thus, $Y\!Z$ theory gives us the following one-loop $n$-gon numerator:
\be
N^{n\text{-gon},YZ}_{(12...n)} = (T^{1}T^{2}\cdots T^{n})+2\, (F^1F^2\cdots F^n)
\ee
where $(\,\cdots)$ indicates an internal contraction over the $YY$ and $\bar{Z}Z$ loops. Plugging in the Feynman rule of \eqn{eq:FeynmanRuleYYZ} and \eqn{eq:FeynmanRuleXZZ}, we can readily obtain expressions in terms of the internal loop factors $l_i$ and the external momenta $k_i$ (we define the $l_i$ loop momentum as that flowing into $k_i$ and out of $k_{i-1}$):
\be
N_{(12...n)}^{\text{NLSM}}=(T^{1}T^{2}\cdots T^{n}) = (l_1 k_1)(l_2 k_2) \cdots (l_n k_n)
\ee
and similarly so for the internal vector contribution:
\be
 (F^1F^2\cdots F^n) = (D-4)(l_1 k_1)(l_2 k_2) \cdots (l_n k_n) + \mathcal{O}(D^0)
\ee
The dimension dependent factor essentially counts that number of internal vector states. While this $n$-gon is manifestly color-dual, it does not produce the right cuts for NLSM. However, the scalar contribution, that comes dressed with an overall factor $(D-4)$ \textit{does} manifest the duality globally, and satisfies all the desired pion cuts. In order for the Feynman rules for $Y\!Z$ theory compute one-loop color-dual numerators consistent with NLSM cuts, we must add some additional states to cancel off he spurious poles, while preserving color-kinematics duality. We leave this as a direction of future work.

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\be
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,l)}-N^{\text{NLSM}}_{(1243,l)}-N^{\text{NLSM}}_{(1342,l)}+N^{\text{NLSM}}_{(1432,l)}
\ee
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $l_i$, we obtain the following expression for the BEL
\be
N^{\text{BEL}}_{1|2,34} = s_{12} (l+k_1)^2 l^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\ee
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\be
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D l}{i\pi^{D/2}} \frac{l^\mu }{l^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\ee
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in large enough dimension, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram.

\section{Two-loop}

@@

\section{Discussion}\label{sec:Discussion}

@@

\bibliographystyle{JHEP}
\bibliography{Refs_2loopNLSM}
\end{document}

>>>>

%Comment commands
\newcommand{\ace}[1]{\textcolor{darkgreen}{\textbf{AE:}{#1}}}

\usepackage[framemethod=default]{mdframed}
\newmdenv[skipabove=7pt,
skipbelow=7pt,
rightline=false,
leftline=false,
topline=false,
bottomline=false,
backgroundcolor=gray!15,
linecolor=gray,
innerleftmargin=5pt,
innerrightmargin=5pt,
innertopmargin=5pt,
innerbottommargin=5pt,
leftmargin=0cm,
rightmargin=0cm,
linewidth=4pt]{eBox}

\tikzset{graviton/.style={decorate, decoration={snake, amplitude=.4mm, segment length=1.5mm, pre length=.5mm, post length=.5mm}, double}}
\tikzfeynmanset{compat=1.1.0}
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\usetikzlibrary{external}  
          %% Load the `external` library
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
  prefix=images/,                    %% Avoid cluttering the directory
 system call={                       %% Use lualatex in system call
    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
  },
]
\tikzfeynmanset{
  fermion2/.style={
    /tikz/postaction={
      /tikz/decoration={
        markings,
        mark=at position 0.7 with {
          \arrow{>[length=6pt, width=7pt]};
        },
      },
      /tikz/decorate=true,
    },
  },
}

@@

\newcommand{\simplebox}{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-1,1){2};
\vertex (a2) at (1,-1){4};
\vertex (a3) at (1,1){3};
\vertex (a4) at (-1,-1){1};
\vertex (mid3) at (.35,.35);
\vertex (mid4) at (.35,-.35);
\vertex (mid5) at (-.35,.35);
\vertex (mid6) at (-.35,-.35);
\diagram{(mid4) --[ultra thick,nhpRed](mid3),
(mid5) --[ultra thick,nhpRed](mid6),
(mid3) --[ultra thick,nhpRed](mid5),
(mid6) --[ultra thick,nhpRed](mid4),
(mid6) --[ultra thick,fermion2](a4),
(mid3) --[ultra thick,fermion2](a3),
(mid4) --[ultra thick,fermion2](a2),
(mid5) --[ultra thick,fermion2](a1),
};
\vertex [dot, scale=1.6](mid3a) at (.35,.35){};
\vertex [dot, scale=1.1,hgrey0](mid3b) at (.35,.35){};
\vertex [dot, scale=1.6](mid4a) at (-.35,.35){};
\vertex [dot, scale=1.1,hgrey0](mid4b) at (-.35,.35){};
\vertex [dot, scale=1.6](mid5a) at (.35,-.35){};
\vertex [dot, scale=1.1,hgrey0](mid5b) at (.35,-.35){};
\vertex [dot, scale=1.6](mid6a) at (-.35,-.35){};
\vertex [dot, scale=1.1,hgrey0](mid6b) at (-.35,-.35){};
\end{feynman}
\end{tikzpicture}
}
}

@@

@@

\author{Alex Edison,}
\author{James Mangan,}
\author{and Nicolas H. Pavao}

\affiliation{Department of Physics and Astronomy, Northwestern
  University, Evanston, Illinois 60208, USA}

\emailAdd{alexander.edison@northwestern.edu, james.mangan@northwestern.edu, pavao@u.northwestern.edu}

\title{\center Towards Globally Color-Dual All-Loop Integrands}

@@

{}_x(F^{a_1}F^{a_2}\cdots F^{a_n})_{y} \equiv F^{a_1}_{\,x|b_2}F^{a_2}_{\,b_2|b_3}\cdots F^{a_n}_{\,b_n|y}

@@

Similarly, we find the following ordered amplitude when plugging in these numerators:

@@

A^{\bar{Z}Z}_{(s,t)} = \frac{n^{\bar{Z}Z}_s}{s_{12}}+\frac{ n^{\bar{Z}Z}_t }{s_{14}} = -s_{13}(p_2+p_3+p_4)^{\mu_1} = s_{13} \,p_1^{\mu_1}

@@

Plugging in the on-shell polarizatoin of the conjugate field in \eqn{eq:onShellZStates} produces precisely the desired result of \eqn{eq:NLSMYZ4point}. Indeed, this construction is valid to all orders at tree-level. One can see this by considering the only two possible factorization channels the contribute the each of these amplitudes, the $YY$ cut and the $\bar{Z}Z$ cut:
\begin{align}
A(...,Y,...,Y,...) &\rightarrow A(...,Y,...,Y)A(Y,...,Y,...)
\\
&\rightarrow A(...,Y,...,Y,...,Z)A(\bar{Z},...)
\end{align}
Both factorization channels are valid descriptions of NLSM amplitudes when plugging in the $Z$ and $\bar{Z}$ on-shell states. As constructed, the $Y$ and $\bar{Z}$ particles may be off-shell, while the $Z$ particles must be placed on shell, $\varepsilon^Z_\mu(k)\rightarrow k_\mu$ in order for these numerators to produce NLSM ordered amplitudes. Given this special property of YZ-theory (namely that the $Z\bar{Z}$ and $YY$ constrctions are equivalent), the $Y$ particle need not be considered when constructing NLSM amplitudes on shell. This is essentially what $J$-theory does, as we will show below.
\paragraph{J-theory} There is an alternative Lagrangian formulation of NLSM constructed by Chueng and one of the authors, that instead starts from the chiral current description of NLSM:

@@

\mathcal{L} = \text{tr}[\partial U \partial U^{-1}] \equiv \text{tr}[J^\mu J_\mu]

@@

where $J_\mu = U\partial_\mu U^{-1}$. To build the J-theory Lagnraigna, we can think of the chiral current $J^\mu$ as a \textit{pure gauge} vector field. Thus, the field strength for $J_\mu$ must vanish, this yielding an equivalent on-shell formulation of the EOM:
\Box J^\mu +g f^{abc} J^\nu J^{\mu\nu} = 0
where we have applied the gauge condition $\partial_\mu J^\mu = 0$. Adding back in a $\bar{J}^\mu$ auxiliary field as a Lagrange multiplier, we obtain the following Lagrangian that is manifestly color-dual:
\begin{eBox}
\mathcal{L}^{\text{J}} = \frac{1}{2}(\partial J)(\partial \bar{J}) +g f^{abc} \bar{J}_{\mu\nu}J^\mu J^\nu
\end{eBox}
where as above, $\bar{J}_{\mu\nu} = \partial_\mu J_\nu - \partial_\nu J_\mu$. Subject to the Lorenz gauge condition, $\partial_\mu J^\mu =0$. This theory is one-to-one with the observables amplitudes of YZ-theory when only considering $Z-\bar{Z}$ on-shell states. As such, the single vertex in this theory is the same as for YZ theory: 
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)
Now we are prepared to discuss the one-loop case. 
\subsection{Color-dual Bootstrap}
Some background here on how we approach color-dual bootstrap. Building an ansatz from irreducible lorentz products (ISPs). Comments about:
\item loop momenta weight counting 
\item polynomial vector spaces 
\item generalized gauge freedom
\item functional jacobi constraints and cuts
\section{Isolating the cubic sector}
Here we'll describe the why constructing the full Yang-Mills kinematic algebra (outside self-dual sector) is generically hard, and why the same reason makes NLSM beyond one-loop hard as well. It has to do with out the vector state sum mixes with the four-point contact. In terms of our previously defined Feynman rules, the three-point YM vertex takes the form:
\cubic{hgrey0}{gluon}{}{gluon}{}{gluon}{}  =  \cubic{hgrey0}{fermion2}{}{}{nhpRed}{}{nhpRed} + \cubic{hgrey0}{}{nhpRed}{fermion2}{}{}{nhpRed} + \Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} +\text{h.c.}
The state-sum between two three-point gluon vertex yields the following expansion in terms of YZ-theory vertices:
 \AcubicB{hgrey0}{gluon}{}{gluon}{}{gluon}{}{gluon}{} =  \AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{fermion2}{} + 5 \text{ perms} + \AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed}
All the $(\epsilon\epsilon)^1$ diagram permutations are isolated to the cubic sector and together manifest color-kinematics duality. However, the exception of the last diagram which introduces a $(\epsilon\epsilon)^2$ factor, must conspire with the Yang-Mills contact to preserve the duality off-shell

\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}{}{nhpRed} = (s_{13}-s_{12}) \epsilon_{(12)} \epsilon_{(34)}
\ee
One was to resolve this is to introduce higher spin modes in order to accomodate the internal index contractions needed to recover the s-channal contract:
\be
\AcubicB{hgrey0}{}{nhpRed}{}{nhpRed}{graviton}{nhpRed}{}{nhpRed} = s_{12 }(\epsilon_{(13)} \epsilon_{(24)} - \epsilon_{(14)} \epsilon_{(23)})
\ee
While introducing succesively higher spiin states to the NMHV lagrangian constructed in \cite{} could in principle work at tree level, we will see that it is not consiisntent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above. This is easily understood in terms of a polarization decomposition introduced by one of the authors \cite{} that decomposes vector amplitudes in terms of building blocks that preserve partial amplitude relations. 
\subsection{Block Decomposition of Yang-Mills}
Describe how different orders in $\varepsilon_i\varepsilon_j$ in pure Yang-Mills captures different cubic and contact sectors of the theory. All the $D$-dimensional Lagnrangians above can be understood as selecting specific on shell states to isolate the cubic sector of Yang-Mills at a given loop order. Start with reducible amplitude block decomp from Ref. \cite:
\begin{eBox}
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\end{eBox}
By a simple mass-dimensioin argument, we know the tree-level cubic sector is simply:
\be
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
and likewise at one-loop:
\be
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\epsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\ee
We can select out both of these manifestly cubic sectors by selecting only $\epsilon^+\epsilon^+$ in light-cone gauge (i.e. SDYM) and one-minus at tree-level OR what $J$-theory and $YZ$-theory select with their on-shell states. Indeed, the $D$-dimensional description of $J$ theory is just the following blocks in the RABD expansion of Yang-Mills:
\be
A(...,\bar{J}_k,...)_{\text{tree}} = \sum_{j\neq k}\Delta^{(jk)}_{\text{tree}}
\ee
and similarly so at one-loop:
\be
A(...)_{\text{1-loop}} \equiv \Delta^{(\varnothing)}_{\text{1-loop}}
\ee
To recover NLSM at one-loop, we need to extract out the $D$-dependent part that corresponds to an internal $YY$-loop from extra-dimensional scalars:
\begin{eBox}
\be
A^{\text{NLSM}}_{\text{1-loop}} \equiv \partial_D \Delta^{(\varnothing)}_{\text{1-loop}}\big|_{\epsilon\rightarrow k}
\ee
\end{eBox}
Why must we take the derivative with respect to $D$? After all, $J$-theory and $YZ$-theory both have well defined propagators for internal $\bar{J}J$ and $\bar{Z}Z$ propagators. However, as we noted in the previous section, all the $J$ and $Z$ states must be on-shell in order to produce NLSM amplitudes. Thus, the unitarity cuts of $J$ theory at one-loop will not produce NLSM amplitudes. However, the internal $YY$-loop is a valid forward limit, since as constructed YZ theory is valid for off-shell Y-particles, and thus they serve as a valid forward limit. We will construct these one-loop integrands below in the next section. 
\subsection{One-loop SDYM Integrands}
\be
\simplebox = g_{\text{YM}}^4 X(k_1,l_1)X(k_2,l_2)X(k_3,l_3)X(k_4,l_4)
\ee
\subsection{One-loop NLSM Integrands}

Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.
\be
\simplebox = (2 f_\pi^{-1})^4 (l_1 k_1)(l_2 k_2)(l_3 k_3)(l_4 k_4)
\ee
where we define sequential loop momenta on the $n$-gon as $l_{i+1}= l_i+k_i$. Given this, the $n$-gon permits a equivalent expression in terms of inverse propagators:
\be
\simplebox =f_\pi^{-4} [12][23][34][41]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_i^L - l_R^2$
\subsection{Two-loop Obstruction}

\section{Bootstrapping two-loop...}

@@

\section{Discussion}\label{sec:Discussion}
\section{... and Beyond}
This could be a section about our expectations for all-loop construction requiring non-local color-dual integrands. Lay out contrafactual proof, i,e. define the following statement:
\be
 Q_L=\textit{``the $L$-loop numerator is polynomial and color-dual"}. 
\ee
Then it is clear that if the kinematic numerator is polynomial and color-dual, evaluating it in kinematic regions where the internal legs are on-shell won't disrupt the kinematic algrebraic relations, and it must still be polynomial. Thus we find:
\be
Q_L \Rightarrow Q_{L-1}
\ee
Or equivalently,
\be
\text{not }Q_{L-1}\Rightarrow \text{not }Q_L
\ee
We have found that the two-loop integrand for Yang-Mills cannot be constructed from a color-dual polynomial kinematic numerator. Thus, if a color-dual representation exists at all-loop order for Yang-Mills it must be a rational function of kinematics.

@@

\appendix
\section{Spinor-Helicity Variables and Conventions}
Dimensions will generically be away from four, in $D=4-2\epsilon$, with $\epsilon$ arbitrary. As noted, it will be convenient to take $D\rightarrow 4$ in certain circumstances after integration.  In such cases we will employ appropriate spinor-helicity variables.  Here we apply the same conventions of Ref.~\cite{jjmcTASI2014}, which we quote now. For massless momenta, $k_a$ and $k_b$, we have 
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]\,,
\end{equation}
the component definition of our spinor bracket that are consistent with the conventions above,
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\end{align}
where the $a_i$ are component values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$. Four-dimensional polarization dot products with fixed helicity states can be mapped as follows:
\begin{equation}\label{eqn:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}\,,
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}\,,
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} \,,
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 \,,
\end{aligned}
\end{equation}
Note that given the above definition, the spinor helicity variables carry the same mass dimension for both the angle and square brackets. In terms of light-cone coordinates that we employed in the text, the angles and squares can be rewritten as follows:
\begin{align}
\langle ab \rangle &= \frac{a_w b_u-a_u b_w}{\sqrt{a_ub_u}}\,,
\\
[ab] &= \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{\sqrt{a_ub_u}}\,,
\end{align}
The conventions we use in the text where the holomorphic spinor cares extra mass dimension essentially amounts to shifting the bracket by the following:
\begin{align}
\langle ab \rangle &\rightarrow a_w b_u-a_u b_w \equiv X(a,b)\,,
\\
[ab] &\rightarrow \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{a_ub_u}\equiv Q(a,b)\,,
\end{align}
With this, we still obtain the same completeness relation for the SDYM spinors:
X(a,b)Q(b,a) = \frac{1}{2}s_{ab}
\section{Regulating BEL integrals}
At one-loop, weight counting tells us that the $n$-gon master numerator must have $n$ on-shell $Z$-particles. Unitarity requires that there are three distinct contributions from $Y\!Z$ theory -- the first from an off-shell $Y$-loop particle and then two more from different orientations of a $\bar{Z}Z$-loop.  Thus, $Y\!Z$ theory gives us the following one-loop $n$-gon numerator:
N^{n\text{-gon},YZ}_{(12...n)} = (T^{1}T^{2}\cdots T^{n})+2\, (F^1F^2\cdots F^n)
where $(\,\cdots)$ indicates an internal contraction over the $YY$ and $\bar{Z}Z$ loops. Plugging in the Feynman rule of \eqn{eq:FeynmanRuleYYZ} and \eqn{eq:FeynmanRuleXZZ}, we can readily obtain expressions in terms of the internal loop factors $l_i$ and the external momenta $k_i$ (we define the $l_i$ loop momentum as that flowing into $k_i$ and out of $k_{i-1}$):
N_{(12...n)}^{\text{NLSM}}=(T^{1}T^{2}\cdots T^{n}) = (l_1 k_1)(l_2 k_2) \cdots (l_n k_n)
and similarly so for the internal vector contribution:
 (F^1F^2\cdots F^n) = (D-4)(l_1 k_1)(l_2 k_2) \cdots (l_n k_n) + \mathcal{O}(D^0)
The dimension dependent factor essentially counts that number of internal vector states. While this $n$-gon is manifestly color-dual, it does not produce the right cuts for NLSM. However, the scalar contribution, that comes dressed with an overall factor $(D-4)$ \textit{does} manifest the duality globally, and satisfies all the desired pion cuts. In order for the Feynman rules for $Y\!Z$ theory compute one-loop color-dual numerators consistent with NLSM cuts, we must add some additional states to cancel off he spurious poles, while preserving color-kinematics duality. We leave this as a direction of future work.
